%-------------------------------%
%  Author: Alessandro Sciarra   %
%    Date: 23 Sep 2020          %
%-------------------------------%

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~%
\begin{frame}[fragile]{How does Bash interpret a line of code?}
    \begin{itemize}
        \item Bash divides each line into words that are demarcated by a \textbf{whitespace character}$^\star$.\tikzmark{whitespace}
        \item \alert{The first word} of the line is the name of \alert{the command} to be executed.
        \item \PP{All the remaining words} become \PP{arguments} to that command (options, filenames, etc.).
    \end{itemize}
    \begin{lstlisting}[style=MyBash, showspaces=true, aboveskip=1mm]
         $@|~|@command arg1 arg2   arg3       arg4
         $@|~|@command arg1 arg2 arg3 arg4
    \end{lstlisting}
    \vspace{2mm}
    \begin{varblock}{alerted}[0.75\textwidth]{}
        \PQ{The amount of whitespace between arguments does not matter!}
    \end{varblock}
    \begin{lstlisting}[style=MyBash, aboveskip=2mm]
        $ echo I am Clark Kent
        |+I am Clark Kent+|                        # <- Same output
        $ echo I    am  Clark         Kent
        |+I am Clark Kent+|                        # <- as here!
    \end{lstlisting}
    \FrameRemark{$^\star$There are a few advanced cases, such as commands that span multiple lines, that have slightly different rules.}
    \begin{tikzpicture}[remember picture, overlay]
        \node[above= 8mm of whitespace, anchor=east, font=\scriptsize, xshift=4pt] (label) {Spaces and Tabs};
        \path[to] (label) edge[out=0, in=0, looseness=1.8] (whitespace);
    \end{tikzpicture}
\end{frame}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~%
\begin{frame}[fragile]{The first hurdle: spaces}{How can such an innocent, invisible character hurt me? ...well...}
    \vspace{-3mm}
    \begin{itemize}
        \item To a shell, whitespace is incredibly important
        \item Don't be fooled into thinking a space or tab more or less won't make much of a difference
        \item  Whitespace is \textbf{vital} to allowing your shell to understand you
    \end{itemize}
    \begin{lstlisting}[style=MyBash, aboveskip=4mm]
        $ ls -1
        |+The secret voice.mp3
        secret+|
        $ rm The secret voice.mp3  # rm gets 3 arguments, not 1!
        |+rm: cannot remove 'The': No such file or directory+|
        |+rm: cannot remove 'voice.mp3': No such file or directory+|
        $ ls
        |+The secret voice.mp3+|       # Where's your file 'secret'!?
    \end{lstlisting}
    \centerline{\ssmall \bash|ls| and \bash|rm| are commands to list the present working directory content and to remove files, respectively.}
    \vspace{-3mm}
    \begin{overlayarea}{\textwidth}{0.3\textheight}
        \begin{varblock}{example}[0.55\textwidth]{}<only@1>
            We will go more deeply into \PS{Word splitting} later!
        \end{varblock}
        \begin{varblock}{}[0.9\textwidth]{Side remark}<only@2>
            \PB{Whitespaces in filenames} should be avoided and \PB{replaced by underscores}. Everything would be much easier. But they are allowed... so deal with it!
        \end{varblock}
    \end{overlayarea}
\end{frame}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~%